\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

This entire project started back in 2015, as a short proposal on a long list of Masters projects, titled ``Ecosystems visualisation and game for scientific outreach". It immediately jumped out at me as my definite first choice, and I was more than happy when I found that it had been allocated to me. I knew at the time that the project would be fun and interesting, but I certainly did not expect that, almost five years later, I would be writing up the journey as a PhD thesis.
I would therefore like to first thank my supervisors, Dan and Samraat, for sticking with me all this way. It has been a pleasure to work with both of you, and I cannot imagine having anybody else guiding me through this path.
% Thank you also to my viva examiners, Daniel and James, for providing excellent feedback and discussions during my five hour viva. 

I would also like to thank all of the gang in the lab, without whom it would have been orders of magnitude more difficult to stay positive through the tough times. Thank you Pamela for expanding and fixing my world view, Juil for always making us laugh, Daphne for your singing and positivity (and memes), Razvan for your kindness and cakes, Sara for your caring nature and comedic genius, Lotte for sacrificing your career as a pop star to bring your positivity to us instead, Jean for being okay at chess but great at weddings, and of course Yang for dealing with me as a housemate for the past few years, even through what has been an arduous 2020 to say the least.
I am grateful to all of you and wish you the very best for the coming years and hope to see you all again.

I must also thank Mum for always being there for me, Dad for inspiring my scientific side, and Jennifer for being an exceptional partner in crime. A warm thank you also goes to Ka Yan, for supporting me and keeping me afloat when work was tough, and when I had lost sight of the light at the end of the tunnel. You mean the world to me.

A huge thank you also goes to fellow PhD student Hsi-Cheng for helping me with the ecological side of EcoBuilder. The game would not look anything like it does now without your help, including the entire genre of the game being puzzle instead of action!
The final release would also not have been possible without help from the members of Neural Reckoning and Pawar Lab, who helped to playtest the game, and volunteered to set up stands at various science festivals to collect player feedback.

Finally, I would also like to thank all the players of EcoBuilder who made the work in the final chapter possible, and for surpassing all expectations I had for how high you would score. It has been a wonderful experience to lead an initiative for crowdsourcing research, and I am proud to have been able to make little citizen scientists out of you all.
My only hope is that you all had fun and learned something along the way, just as I did on my journey towards completing this thesis.

\end{acknowledgements}