\chapter{Introduction}
choice 1: Human understanding is a complex process, and conveying knowledge to others is one of the most important tasks available to us, and the primary highway into understanding is through our eyes.
choice 2: Social networks, metabolic pathways, the internet. Networks exist all around us and are becoming an incredibly important data structure. At the time of writing, networks are being...
choice 3: At the time of writing, the video game industry just surpassed film as the most profitable mode of entertainment in the world. And it is only getting higher.

% """
%  From the 1980s, industrial demand for graph drawing algorithms has grown
% – Software engineering: CASE systems, reverse engineering – Biology: PPI networks, gene regulatory networks
% – Physical networks: network management tools
% – Security: risk management, money movements
% – Social network analysis
% – Customer relationship management: value identification Many companies buy graph drawing algorithms, many code them.
% Currently the international market for graph drawing algorithms is in the hundreds of millions of dollars per year.
% """

The background literature review is split up into the beginning of the three main chapters. It is designed to be read in one swoop.

\section{Motivation and Objectives}

Motivation and Objectives here.
some literature stuff too?

\section{Contributions}

Contributions here.
Algorithms for SGD, HEB, PCD.


\section{Statement of Originality}
This is to certify that to the best of my knowledge, the content of this thesis is my own work. This thesis has not been submitted for any other degree or other purposes.

I certify that the intellectual content of this thesis is the product of my own work and that all the assistance received in preparing this thesis and sources have been acknowledged.

\section{Publications}

The work presented in Chapter~\ref{sec:sgd} was previously published in Zheng et al.\ \cite{Zheng2019Stochastic}.
An early version of the work presented in Section~\ref{sec:heb_background} was previously published in Zheng et al.\ \cite{Zheng2018}
The work presented in Section~\ref{sec:power_graph} has previously been published in Zheng et al.\ \cite{Zheng2019Power}.