\chapter{EcoBuilder}

The story behind EcoBuilder starts with a howl---the howl of wolves in Yellowstone park in America. In the years leading up to 1995, ecologists noticed that the ecosystem in Yellowstone was not as healthy as it used to be. In 1926, wolves were declared officially extinct in the park due to a lack of anti-hunting regulation at the time. The loss of the wolves, the park's top predator, meant that elk no longer had any predator, leaving them free to eat up as many plants as they wanted without risk of predation. This led to the abundance of plants hitting an all time low, causing smaller animals such as beavers and fish to struggle and eventually go extinct as well.
Attempts were made by wildlife conservationalists to suppress the elk population by moving or hunting them, but were not sufficient to bring the ecosystem back to its original state. Ecologists then came up with the idea of reintroducing wolves to try and save the ecosystem, and in January of 1995 

After watching a documentary about this, Dr. Dan Goodman at Imperial College came up with an idea: wouldn't it be cool to be able to do this yourself? 

And so he set out to make a game based on this. Dr. Goodman is a neuroscientist, however, and so to supplement the ecology side he contacted Dr. Samraat Pawar in the Life Sciences department and together proposed a student project to make the game. That is where I come in, as a PhD student looking for a cool project. 3 years later, and we are finally ready to release the final version of the game.

EcoBuilder is a puzzle game where you will build your own ecosystem of plants and animals, modelled using the same simulations that researchers use to study real ecosystems. 

Outreach and teaching is a primary part of the game, and raising awareness for environmental issues is as important as ever. Just think of ecosystems collaspsing etc
You will learn about various ecological phenomena along the way

The game also has a citizen science element to it, where if you get really good at playing then your high scores will help real researchers to understand ecosystems better!

\section{Background}
application of SGD in an interactive setting

\subsection{Structure}
trophic levels
chain length
(mention loops)
\subsection{Dynamics}
May stability
feasibility

\section{Design and Hypotheses}
\subsection{Ap}
score
trophic levels

layout:
\begin{itemize}
    \item TROPHIC LEVELS:
    \begin{itemize}
        \item gauss-seidel iteration because positive definite, optimised to only take O(m) per iteration and O(n) extra space (because we only care about the sum of the rows and not where the rows are placed. In general we find the number of iterations required is <100 (test this) but requires for loops~\cite{oliviasimpsonpaper}
        \item whether this laplacian has a solution can also be found using a breadth first search from sources (which is also used for chain)
        \item note that this is the same set of equations as in \eqref{tutte}
    \end{itemize}
    \item LAYOUT:
    \begin{itemize}
        \item SGD is so fast that we use it on every topological change
        \item 3D vs 2D: originally used 3D, but we found that players could not quickly pick up the rotational interface. Nodes would also often become obscured in the center of the layout, so we switched to 2D instead. 
        \item however this also made trophic level constraints more difficult to satisfy because there is one fewer dimension to move around in (previously all that was needed was to always keep the y-position. The best solution was to add 5 iterations of $\mu=1$ at the beginning, as an initialization step.
        \item localised majorization is used every frame for one node in order to fine-tune layout
        \item we compare constrained to non-constrained to see who performs better. They both use the same initial layout (SGD constrained to trophic levels) but with either constrained or non-constrained majorization when fine tuning.
        \item TODO: procrustes then performed on each connected component
    \end{itemize}
    
    \item superfocus
    \item johnson's algorithm for loops
    \item mention old idea for chess-board and why it didn't work
\end{itemize}

model:
\begin{itemize}
    \item scoring metric
    \item diagonals of interaction matrix (reference hsi-cheng's work)
    \item math.net library
\end{itemize}

UI:
\begin{itemize}
    \item database in sql
    \item indexed using...
\end{itemize}


\section{Results and Analysis}
